---
title: "Disposition screening"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Disposition screening}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# Auth/network chunks are shown but not run at build time; the disposition_rollup
# chunk below runs, so its output is real.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

Screen a phone sample against the **disposition dataset** -- one row per `(phone, campaign_id)` for every recipient Survey160 has contacted across closed campaigns -- to drop numbers already completed or refused before you field. Assumes you have completed First-time setup (see the package README): GCS access to `s160_disposition_prod`.

## Screen a sample

Pull the dataset once (cached), then annotate your sample in place:

```{r}
library(survey160r)
s160_gcs_init("s160_disposition_prod")   # one browser sign-in, cached
dataset <- disposition_pull()            # ~140 MB first pull, then cached

cleaned <- disposition_screen(my_sample, dataset)
kept <- subset(cleaned, !(ever_complete %in% TRUE | ever_terminated %in% TRUE))
```

`disposition_screen()` returns `my_sample` unchanged with the screening columns appended 1:1 (see [What you get back](#what-you-get-back)). The `%in% TRUE` filter drops completed and refused numbers; blank or unparseable phones screen to all-`NA` and are kept, so they surface for correction rather than vanishing.

## Look up specific numbers

`disposition_summary()` runs the per-phone rollup as a standalone query; `disposition_records()` returns the raw stored rows, one per `(phone, campaign_id)`:

```{r}
disposition_summary(dataset, phones = c("2015550101", "2015550102"))
disposition_records(dataset, campaign_ids = 2339)
```

Both take `phones` / `campaign_ids` filters. Phones match digit-normalized: a leading US `1` is dropped, so an 11-digit number matches a stored 10-digit one.

## Screen several samples

Each reader call reads the whole file, so read once and roll up in memory per sample. `disposition_rollup()` is pure -- here it is on an in-memory frame, which runs:

```{r rollup, eval = TRUE}
library(survey160r)
records <- data.frame(
  phone          = c("5551234567", "5551234567", "5559876543"),
  campaign_id    = c(101L, 102L, 101L),
  engaged        = c(1L, 1L, 0L), opt_in       = c(1L, 0L, 0L),
  complete       = c(1L, 0L, 0L), web_complete = c(0L, 0L, 0L),
  terminated     = c(0L, 1L, 0L),
  date_closed_on = as.Date(c("2026-01-10", "2026-01-20", "2026-01-15"))
)
disposition_rollup(records, phones = c("5551234567", "5550000000"))
```

In your own code, read once and reuse: `records <- disposition_records(dataset)`, then `disposition_rollup(records, phones = ...)` for each list. (The file readers take a `dataset` path; `disposition_rollup()` takes the in-memory `data` frame that `disposition_records()` returns.)

## What you get back

`disposition_screen()` and `disposition_summary()` return one row per phone:

| Column | Meaning |
|---|---|
| `ever_contacted` | Ever sent an intro (present in the dataset). |
| `n_campaigns` | Distinct campaigns that contacted it. |
| `ever_engaged` | Ever replied. |
| `ever_opted_in` | Ever consented. |
| `ever_complete` | Ever completed (reached the SMS close, or a web completion). |
| `ever_terminated` | Ever screened out or refused. |
| `latest_disposition` | Outcome of the most recent campaign (funnel below). |
| `campaigns` | Comma-separated campaign ids. |

`latest_disposition`, least to most advanced: `never_contacted` -> `non_response` -> `engaged` -> `opt_in` -> `terminated` -> `complete` -> `web_complete`. `disposition_records()` returns the stored per-campaign columns instead; see `?disposition_records`.

## Notes

- **Contacted-only.** A loaded number that was never texted is not in the dataset; it screens as `never_contacted`.
- **Text-to-web (external) campaigns** have no completion callback, so `complete` is `NA` and never a false positive.
- **`date_closed_on` and `error` are `NA` in the beta**, so the `date_from` / `date_to` filters return no rows.
- **Stale cache.** The dataset is re-published periodically; `disposition_pull(refresh = TRUE)` forces a fresh download.
- **`env`.** `disposition_pull()`'s `env` is `"prod"` (default) or `"dev"` -- the disposition buckets have no staging tier, so this differs from `s160_api_auth()`'s `"prod"` / `"staging"`.

## See also

`?disposition_screen`, `?disposition_summary`, `?disposition_records`, `?disposition_pull`. To build the dataset from a raw campaign CSV (the producer side), `?disposition_run` and `?disposition_input_columns`. For latency reports, `vignette("latency")`.
