---
title: "Disposition screening"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Disposition screening}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# Auth/network chunks are shown but not run at build time; the disposition_summary
# chunk below (on an in-memory frame) runs, so its output is real.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

Screen a phone sample against the **disposition dataset** -- one row per `(phone, campaign_id)` for every recipient Survey160 has contacted across closed campaigns -- to drop numbers already completed, refused, or screened out before you field. Assumes you have completed First-time setup (see the package README): GCS access to `s160_disposition_prod`.

## Screen a sample

Pull the dataset once (cached), then annotate your sample in place:

```{r}
library(survey160r)
s160_gcs_init()   # one browser sign-in, cached
dataset <- disposition_pull()            # a few hundred MB first pull, then cached

my_sample <- data.frame(phone = c("2015550101", "2015550102"))   # your phone sample
cleaned <- disposition_screen(my_sample, dataset)

# keep numbers not already finished: not completed (SMS or web), and not a hard
# stop (declined, screened out, or an unsplit terminal)
kept <- subset(
  cleaned,
  !((n_completed > 0 | n_web_complete > 0) %in% TRUE |
      (n_terminated > 0 | n_refused > 0 | n_ineligible > 0) %in% TRUE)
)
```

`disposition_screen()` returns `my_sample` unchanged with the screening columns appended 1:1 (see [What you get back](#what-you-get-back)). The columns are cumulative **counts** across the number's campaigns, so `> 0` means "reached that status at least once". `n_terminated`, `n_refused` (declined), and `n_ineligible` (screened out) are the hard-stop signals; on the prod projection they are set independently -- a number can be flagged `refused` / `ineligible` without a bare `terminated`, and vice versa -- so screen on all three. The `%in% TRUE` guard keeps blank or unparseable phones (they screen to all-`NA`), so they surface for correction rather than vanishing.

## Look up specific numbers

`disposition_summary()` runs the per-phone rollup as a standalone query; `disposition_records()` returns the raw stored rows, one per `(phone, campaign_id)`:

```{r}
disposition_summary(dataset, phones = c("2015550101", "2015550102"))
disposition_records(dataset, campaign_ids = 1234)
```

Both take `phones` / `campaign_ids` filters. Phones match digit-normalized: a leading US `1` is dropped, so an 11-digit number matches a stored 10-digit one. `disposition_records()` hands back the stored schema as-is, one row per `(phone, campaign_id)` -- the funnel flags, the recipient's carrier and any delivery error, the Tracker dimensions, and the disposition date. See [The stored record columns](#the-stored-record-columns).

## Screen several samples

Each reader call reads the whole file (nanoparquet has no predicate pushdown), so read once and summarize in memory per sample. `disposition_summary()` accepts either the dataset path or an already-read frame; on a frame it does no I/O, so this runs:

```{r summary_frame, eval = TRUE}
library(survey160r)
records <- data.frame(
  phone          = c("5551234567", "5551234567", "5559876543"),
  campaign_id    = c(101L, 102L, 101L),
  engaged        = c(1L, 1L, 0L), opted_in     = c(1L, 0L, 0L),
  completed      = c(1L, 0L, 0L), web_complete = c(0L, 0L, 0L),
  refused        = c(0L, 1L, 0L), ineligible   = c(0L, 0L, 0L),
  terminated     = c(0L, 1L, 0L),
  disposition_date = as.Date(c("2026-01-10", "2026-01-20", "2026-01-15"))
)
disposition_summary(records, phones = c("5551234567", "5550000000"))
```

Note `5551234567`: its **latest** disposition is `refused` (its most recent campaign, 102, ended in a refusal) while its **best** is `completed` (campaign 101 finished the survey). In your own code, read once and reuse: `records <- disposition_records(dataset)`, then `disposition_summary(records, phones = ...)` for each list.

## What you get back

`disposition_screen()` and `disposition_summary()` return one row per phone:

| Column | Meaning |
|---|---|
| `phone` | The digit-normalized number. |
| `n_campaigns` / `campaigns` | How many campaigns contacted it (`0` = never contacted), and the comma-separated id list. |
| `n_engaged` | Campaigns where it replied. |
| `n_opted_in` | Campaigns where it consented. |
| `n_completed` / `n_web_complete` | Campaigns it completed on SMS (reached the close) / on the web. |
| `n_terminated` | Campaigns the producer flagged as a hard stop (the `terminated` flag). |
| `n_refused` / `n_ineligible` | Campaigns where the terminal was a decline / a screen-out (name-classified). On the prod projection these are set independently of `n_terminated` and need not sum to it -- so screen on all three. |
| `n_error` | Campaigns that carried a carrier delivery-error code. |
| `latest_disposition` / `latest_campaign_id` | Outcome of the **most-recent** campaign, and that campaign's id. |
| `best_disposition` / `best_campaign_id` | **Furthest-reached** outcome across all its campaigns, and the id that reached it. |
| `first_disposition_date` / `last_disposition_date` | Earliest and latest disposition date across its campaigns. |

The counts overlap rather than partition `n_campaigns` -- a completed campaign is also engaged -- so read each as "reached status X in this many campaigns".

`latest_disposition` and `best_disposition`, least to most advanced: `never_contacted` -> `non_response` -> `engaged` -> `opted_in` -> `terminated` -> `ineligible` -> `refused` -> `completed` -> `web_complete`. The terminal band (`terminated` < `ineligible` < `refused`) sits between `opted_in` and `completed`: `terminated` is the unsplit residual, `ineligible` a screen-out, `refused` a decline. `disposition_records()` returns the stored per-campaign columns instead; see [The stored record columns](#the-stored-record-columns).

## The stored record columns

`disposition_records()` returns the dataset as stored -- one row per `(phone, campaign_id)`, contacted-only -- which `disposition_summary()` and `disposition_screen()` roll up per phone (above). The columns, in funnel order:

| Column | Meaning |
|---|---|
| `phone` / `campaign_id` | The row key: the digit-normalized number and the campaign that contacted it. |
| `sent` | Always `1` -- the dataset holds only contacted numbers. |
| `engaged` | Replied at all. |
| `opted_in` | Consented to take the survey. |
| `completed` | Reached the survey close in-channel. `NA` on text-to-web external (no completion callback). |
| `web_complete` | Completed on the web. Equals `completed` for a `t2w` campaign; `0` for `sms` and `t2w_external`. |
| `terminated` | Hard-stop flag. On the prod projection it is set independently of the split below -- a row can be `terminated` with neither `refused` nor `ineligible`, and either can be set without `terminated` -- so screen on all three. |
| `refused` / `ineligible` | The name-classified terminal split -- a decline / a screen-out. |
| `error` | The raw carrier delivery-error code for the send; `NA` when the export carries no usable code. |
| `carrier` | The recipient's mobile network, from the export's optional `misc` column; `NA` when that column is omitted or blank. |
| `loi` / `topic` | Length of interview in minutes, and the campaign topic, from the Project Tracker. |
| `mode` | Send channel: `sms`, `t2w`, or `t2w_external`. |
| `registration_id` | The campaign's 10DLC registration id, from the Project Tracker. |
| `disposition_date` | The number's last-send date, or `NA` when no send survives archival. |

The funnel flags overlap rather than partition -- a `completed` row is also `engaged` -- so read each as a status the row reached, not an exclusive bucket. Provenance of the columns: `sent`, the funnel flags, `mode`, `error`, and `carrier` are derived from the campaign export; `loi`, `topic`, and `registration_id` come from the Project Tracker, so they are populated only for campaigns the Tracker covers and read `NA` on older campaigns that predate it. An un-enriched projection (published before a column existed, or without Tracker enrichment) simply omits that column, and `disposition_records()` returns whatever subset is present -- see `?disposition_records`.

## Notes

- **Contacted-only.** A loaded number that was never texted is not in the dataset; it screens as `never_contacted`.
- **Text-to-web (external) campaigns** have no completion callback, so `completed` is `NA` -- never a false positive, but an external completer reads as `n_completed = 0` and is kept (a false negative you cannot screen out here).
- **`disposition_date` and `error` are populated.** The live producer stamps each phone's last-send date and any carrier delivery-error code, so the `date_from` / `date_to` filters and `n_error` work (prod currently spans mid-2023 to the last pipeline pass). A row with no surviving send has an `NA` `disposition_date`, which any date bound drops.
- **Terminal split rollout.** `refused` / `ineligible` are set for the terminals the routing could name-classify; a small residual stays `terminated`. A projection produced before the split defaults them to `0`, so those rows read as plain `terminated`.
- **Screen on the counts, not on `latest_disposition`.** `latest_disposition` / `best_disposition` are informational; a number whose most-recent campaign only reached `engaged` can still have completed an earlier one, so filter on the `n_*` counts (which see every campaign), not the latest label.
- **Scale.** The projection is tens of millions of rows and a few hundred MB; each reader reads it whole, so pull once and summarize per sample in memory (see [Screen several samples](#screen-several-samples)).
- **Stale cache.** The dataset is re-published periodically; `disposition_pull(refresh = TRUE)` forces a fresh download.
- **`env`.** `disposition_pull()` reads `"prod"` (default) or `"dev"` -- there is no `staging` tier (staging sends fold into the prod bucket), so `env = "staging"` errors; use `env = "prod"` for staging data. Run `s160_datasets()` to see every dataset's environments.

## See also

`?disposition_screen`, `?disposition_summary`, `?disposition_records`, `?disposition_pull`. To build the dataset from a raw campaign CSV (the producer side), `?disposition_run` and `?disposition_input_columns`. For latency reports, `vignette("latency")`.
