---
title: "Disposition screening"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Disposition screening}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# The GCS / whole-dataset examples need auth and the ~137 MB projection, so they
# are shown but not executed at build time. The pure-core example below (rolling
# an in-memory frame up) does run, so its output is real.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

**The Survey-Manager surface.** Screen a phone sample against the **disposition dataset** -- a shared, phone-indexed record of every recipient Survey160 has contacted across closed campaigns (one row per `(phone, campaign_id)`, contacted-only) -- to answer "which of these numbers have we contacted / completed / refused before?" and clean a sample list before you field it.

> Disposition screening needs a recent survey160r (these functions arrived in 0.24.0). R-universe rebuilds from `main` and can lag a release by up to ~30 minutes; if `disposition_pull()` isn't found after installing, get the latest straight from GitHub with `pak::pkg_install("survey160/survey160r")`.

## Pull once, screen a sample in place

```{r}
library(survey160r)

# 1. Authenticate to GCS once (browser sign-in on first run; token cached)
s160_gcs_init(bucket = "s160_disposition_prod")

# 2. Download the shared dataset to a local, cached path
dataset <- disposition_pull()          # env = "prod" (default) or "dev"

# 3. Annotate your sample in place -- your original rows and columns are kept,
#    the screening columns are appended 1:1
cleaned <- disposition_screen(my_sample, dataset, phone_col = "phone")

# 4. Drop the numbers you've already finished with, then write the list out.
#    Blank/unparseable phones come back all-NA and are KEPT (so they surface
#    for correction rather than silently vanishing).
kept <- subset(cleaned, !(ever_complete %in% TRUE | ever_terminated %in% TRUE))
write.csv(kept, "sample_screened.csv", row.names = FALSE)
```

For example, screening a small list that carries its own strata columns (output illustrative):

```{r}
my_sample <- data.frame(
  phone  = c("2015550101", "2015550102", "9999999999"),
  region = c("NE", "NE", "SW"),
  quota  = c("A", "A", "B")
)
disposition_screen(my_sample, dataset)
```

which returns your sample with the eight screening columns appended (output illustrative):

```
       phone region quota ever_contacted n_campaigns ever_engaged ever_opted_in ever_complete ever_terminated latest_disposition campaigns
1 2015550101     NE     A           TRUE           2         TRUE          TRUE          TRUE           FALSE           complete 2339,2354
2 2015550102     NE     A           TRUE           1         TRUE         FALSE         FALSE            TRUE         terminated      2339
3 9999999999     SW     B          FALSE           0        FALSE         FALSE         FALSE           FALSE    never_contacted      <NA>
```

Your `region` / `quota` columns come back untouched and the screening columns are appended 1:1, so the same `!(ever_complete %in% TRUE | ever_terminated %in% TRUE)` filter keeps only the never-contacted number here -- the already-completed and refused rows drop out before you field (and an all-`NA` blank-phone row would stay, for correction).

## Columns `disposition_screen()` appends

Each is computed across every campaign that contacted the number:

| Column | Meaning |
|---|---|
| `ever_contacted` | `TRUE` if the number was ever sent an intro (i.e. present in the dataset). |
| `n_campaigns` | How many distinct campaigns contacted it. |
| `ever_engaged` | Ever replied to an intro. |
| `ever_opted_in` | Ever gave consent (accepted the opt-in question). |
| `ever_complete` | Ever completed a survey (reached the SMS close, or a web completion). |
| `ever_terminated` | Ever screened out (ineligible) or refused. |
| `latest_disposition` | The outcome of the number's most recent campaign (see below). |
| `campaigns` | Comma-separated campaign ids that contacted the number. |

`latest_disposition` is one value, in funnel order:
`never_contacted` (not in the dataset) -> `non_response` (contacted, no reply) -> `engaged` (replied) -> `opt_in` (consented) -> `terminated` (screened out / refused) -> `complete` -> `web_complete`.

## The pure core

`disposition_screen()` and `disposition_summary()` are thin readers over a pure rollup, `disposition_rollup()`, which takes an in-memory frame (one row per `(phone, campaign_id)`) and returns one row per phone. It has no I/O, so it runs anywhere -- here on a hand-built frame (this block actually executes when the vignette is built):

```{r rollup, eval = TRUE}
library(survey160r)
records <- data.frame(
  phone          = c("5551234567", "5551234567", "5559876543"),
  campaign_id    = c(101L, 102L, 101L),
  engaged        = c(1L, 1L, 0L),
  opt_in         = c(1L, 0L, 0L),
  complete       = c(1L, 0L, 0L),
  web_complete   = c(0L, 0L, 0L),
  terminated     = c(0L, 1L, 0L),
  date_closed_on = as.Date(c("2026-01-10", "2026-01-20", "2026-01-15"))
)
disposition_rollup(records, phones = c("5551234567", "5550000000"))
```

The first number rolls up across both its campaigns (`n_campaigns = 2`, latest disposition from the newer close date); the screened-but-absent `5550000000` comes back `never_contacted`.

## Ad-hoc queries over the same dataset

```{r}
# One row per phone: the cross-campaign screening flags + latest_disposition
disposition_summary(dataset, phones = c("5551234567", "5559876543"))

# The raw rows beneath that rollup: one per (phone, campaign_id)
disposition_records(dataset, campaign_ids = 1234)
```

**Screening several samples? Read once.** Every reader call above loads the whole dataset into memory (there is no server-side filtering), so re-reading it is the main cost. Read the records once, then roll up in memory per sample -- no further I/O:

```{r}
records <- disposition_records(dataset)                 # read the file once
disposition_rollup(records, phones = sample_a$phone)    # pure, in-memory
disposition_rollup(records, phones = sample_b$phone)
```

Phone numbers are matched digit-normalized (a leading US `1` is dropped, so an 11-digit number matches a stored 10-digit one), and the original formatting in your sample is preserved.

## Beta caveats

- **Contacted-only.** A loaded number that was never texted is not in the dataset; screening returns it as `never_contacted`.
- **Text-to-web (external) campaigns.** Completion happens on an external platform with no callback, so `complete` is unknowable and left `NA`; such a row falls back to its last known in-channel step and is never reported as a false `complete`.
- **`date_closed_on` and `error` are `NA` in the beta**, so the `date_from` / `date_to` filters return no rows for now.

## Producer side

Building the per-respondent disposition frame from a raw campaign CSV is `disposition_run(campaign_id, data)` -- pure and source-agnostic, like `latency_run()`. `disposition_input_columns()` gives the column-projection set (mirroring `latency_input_columns()`). Persisting the frame as Parquet, walking the fleet, and publishing the shared dataset live in the downstream consumer project.
