---
title: "Latency analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Latency analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# Examples read a real campaign CSV / need GCS auth, so they are shown but not
# run at build time.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

Turn a campaign's raw results CSV into a per-campaign recipient-latency report -- how fast each message followed the previous one, and what share of recipients hit the fleet's target windows -- so you can tune send timing. Assumes you have completed First-time setup (see the package README) if you read from GCS.

## Build a report

Read the campaign's results, then run the report:

```{r}
library(survey160r)
s160_gcs_init()            # one browser sign-in, cached
data   <- s160_gcs_campaign_results_read(1234, hash = TRUE)  # or s160_read_csv("path/to/1234.csv")
result <- latency_run(1234, data, field_timezone = "America/New_York")
```

You already read these CSVs in your own scripts; `s160_gcs_campaign_results_read(hash = TRUE)` and `s160_read_csv()` just add the source-hash provenance that `latency_run()` records on `result$meta`. `field_timezone` sets the local time zone for the hour-of-day buckets. `latency_run()` builds the config from the CSV header and returns a list: `consolidated` (the report), `latency_frame` (per-respondent detail), `diagnostics` (parse / NA counts), and `meta` (provenance).

## Customize the config

`latency_run()` derives the config from the CSV header; override through `...`, or build it explicitly to inspect first, then validate before running:

```{r}
config <- latency_build_config(1234, data,
                               field_timezone = "America/New_York",
                               project_id     = 9999,
                               date_filter    = "2026-01-26")   # keep a single field date
latency_validate_config(config, data)   # fail fast: required columns, flow order, no terminal states
result <- latency_report(data, config)
```

## Run deterministically

`latency_report()` is the pure entry point (no I/O). Pass `run_at` to fix `run_at_utc` -- otherwise it comes from the clock -- for reproducible output and tests:

```{r}
latency_report(data, config,
               run_at = as.POSIXct("2026-01-27 12:00:00", tz = "UTC"))
```

A hand-built data frame works too; `result$meta` provenance is then `NA`, which is correct for that case.

## What you get back

`result$consolidated` is one row per `(campaign_id, date, hour_local, segment, threshold_min)` -- the shape a consumer project writes to Parquet:

| Column(s) | Purpose |
|---|---|
| `campaign_id`, `project_id` | Wave identity |
| `survey_mode` | `sms` / `t2w` / `t2w_external` |
| `date`, `hour_local` | Bucket grain (`hour_local` is `NA` for the day rollup) |
| `segment`, `segment_index` | Pair of consecutive flow questions, e.g. `intro->q1` |
| `threshold_min` | Fleet threshold (1, 3, 5, or 10 min) |
| `n`, `pct_le` | Per-segment in-window dispatch metrics |
| `n_respondents`, `pct_resp_hit_gt`, `pct_resp_worst_gt` | Respondent-cascade metrics |
| `mean_delta_min`, `p50_delta_min`, `p90_delta_min`, `p95_delta_min` | Per-segment latency distribution (minutes) |
| `n_na_parse`, `n_na_missing`, `n_na_chain` | Segments with no computable latency, by reason |
| `n_sent`, `n_engaged`, `n_opted_in`, `n_completed`, `n_ineligible` | Bucket volume: sent, engaged, opted-in, completed, screened out |
| `algorithm_version`, `config_hash`, `source_csv_hash`, `run_at_utc`, `run_by` | Provenance |

`result$meta` carries `algorithm_version`, `schema_version`, `config_hash`, `run_at_utc`, `source_csv_hash`, and `source_csv_path`. See `?latency_run` for the full argument list.

## Notes

- `latency_run()` composes `latency_build_config()` + `latency_report()`; reach for the pieces only when you need to inspect or override the config.
- `respondent_id_column` defaults to `NULL` -- `userid` is the agent login, not a per-respondent id.

## See also

`?latency_run`, `?latency_report`, `?latency_build_config`, `?latency_validate_config`. For disposition screening, `vignette("disposition")`.
