---
title: "Latency analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Latency analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# Examples read a real campaign CSV / need GCS auth, so they are shown but not
# executed at build time (like the package's `\dontrun` help examples).
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

When a wave goes out you want to know how quickly it actually reached people: how fast each message followed the one before it, and what share of recipients hit the fleet's target windows. This guide walks through building that report for one campaign and reading the numbers it returns.

survey160r is *algorithm-only*: you hand it a campaign's results as a data frame and it returns the report as an in-memory R object. It does not fetch, write, or schedule anything -- so the first question is where your data comes from.

## Step 1 -- get the data

You need one campaign's raw results as a data frame. Two readers produce one with the provenance attributes the report expects; pick whichever matches where the CSV lives.

```{r}
library(survey160r)
s160_gcs_init(bucket = "campaign_results")   # one-time browser sign-in, then cached

# From the last GCS export...
data <- s160_gcs_pull_csv(1234)

# ...or from a local file / Dropbox archive:
data <- s160_read_csv("~/Dropbox/archive/campaign_1234.csv")
```

Both readers stamp the source hash and canonical path onto the frame, and the report surfaces them on `result$meta` -- so every run is traceable to its exact input. (A hand-built data frame works too; the provenance just comes back `NA`, which is correct for that case.)

## Step 2 -- run the report

Now run it. `latency_run()` builds the config from the CSV header and runs the algorithm in one call:

```{r}
result <- latency_run(1234, data, field_timezone = "America/New_York")
```

`field_timezone` is the campaign's local time zone -- it is what the hour-of-day buckets are computed in. The result is a list with four pieces:

```{r}
names(result)
#> [1] "consolidated"  "latency_frame" "diagnostics"   "meta"
```

- **`consolidated`** -- the report itself: metrics per `(date, hour, segment, threshold)`. This is the table you will usually work with.
- **`latency_frame`** -- one row per `(respondent, segment)`, the per-respondent detail the consolidated table aggregates, each tagged with why it is `NA` when a latency could not be computed.
- **`diagnostics`** -- counts of parse failures, non-computable segments by reason, clamped negatives, and a respondent summary.
- **`meta`** -- provenance: `algorithm_version`, `schema_version`, `config_hash`, `run_at_utc`, and the source hash/path.

## Step 3 -- read the numbers

Look at `result$consolidated`. Each row is one `(campaign_id, date, hour_local, segment, threshold_min)` cell:

```{r}
head(result$consolidated)
```

The columns fall into a few groups:

| Column(s) | Purpose |
|---|---|
| `campaign_id`, `project_id` | Wave identity |
| `survey_mode` | `sms` / `t2w` / `t2w_external` |
| `date`, `hour_local` | Bucket grain (`hour_local` is `NA` for the day rollup) |
| `segment`, `segment_index` | Pair of consecutive flow questions, e.g. `intro->q1` |
| `threshold_min` | Universal fleet threshold (1, 3, 5, or 10 min) |
| `n`, `pct_le` | Per-segment in-window dispatch metrics |
| `n_respondents`, `pct_resp_hit_gt`, `pct_resp_worst_gt` | Respondent-cascade metrics |
| `mean_delta_min`, `p50_delta_min`, `p90_delta_min`, `p95_delta_min` | Per-segment latency distribution (minutes) |
| `n_na_parse`, `n_na_missing`, `n_na_chain` | Segments with no computable latency, by reason |
| `n_texted`, `n_engaged`, `n_consented`, `n_completed`, `n_ineligible` | Campaign volume for the `(campaign, date, hour)` bucket -- texted (sent to), engaged (replied), consented, completed, screened out -- repeated across that bucket's segment/threshold rows |
| `algorithm_version`, `config_hash`, `source_csv_hash`, `run_at_utc`, `run_by` | Provenance |

This is exactly the column shape a downstream project writes to Parquet: each row stands on its own, so the table needs no sidecar manifest.

## Customizing the run

`latency_run()`'s config is derived from the CSV header, but you can override it. Under the hood it calls `latency_build_config()`, which you can also call directly to inspect or tweak the config before running:

```{r}
config <- latency_build_config(
  campaign_id = 1234,
  data = data,
  field_timezone = "America/New_York",
  project_id = 9999,
  date_filter = "2026-01-26",      # keep a single field date
  respondent_id_column = NULL      # `userid` is the agent login, not per-respondent
)
```

`latency_report(data, config)` then runs the pure algorithm on that config. Passing a `run_at` makes the whole run deterministic (otherwise `run_at_utc` comes from the clock) -- handy for tests and reproducible output:

```{r}
result <- latency_report(data, config,
                         run_at = as.POSIXct("2026-01-27 12:00:00", tz = "UTC"))
```

`latency_report()` is the pure entry point -- no I/O, no mutable globals -- so it is also the one to reach for in tests and ad-hoc analysis.

## Checking the config

Before a run, `latency_validate_config()` fails fast on the common mistakes -- a missing required column, a flow order that does not match the data, an unknown key, or a terminal state (`refusal` / `ineligible`) listed among the flow questions:

```{r}
latency_validate_config(config, data)
```

Once the report is in hand, persisting it as Parquet, walking the whole campaign fleet, and scheduling all live in the downstream consumer project -- survey160r's job ends at the in-memory result.
